Making (Secure) FETCH Happen: \\ \vspace{2 mm} {\large End-to-End Metadata Security for IMAP}

### 2015-04-22 paper.tex
mixing

If a message has been received at the client from a secure source as in Section~\ref{mail-from-secure}, we would like to upload it to the IMAP server so that other \project clients can view it, and to achieve user-facing functionality backwards compatibility. Yet if we push a message to the server immediately upon its receipt, we are leaking the time of message receipt to the server. Thus, we need additional mechanisms for ensuring that a server that stores messages cannot determine when a message arrived.

@@

  \draw [arrow] (\project.10) -- node [above] {State information} (scheduler.-190);
  \draw [arrow] (scheduler.190) -- node [below] {Next action: UP/DOWN} (\project.-10);

@@

This scheduler receives information from the \project module about its state. Namely, it is informed when a message is received at the client or marked for deletion on the server. Instead of immediately performing this action, the \project module will wait until the next timer tick, when it is called to perform its next scheduled action by its client. At that time, it asks the scheduler whether its next action is an UP or DOWN action. The scheduler considers how often messages are being received and how often scheduled communications happen, and returns an UP or DOWN result to keep overall server growth consistent with actual message receipt and deletion rates. The server will be able to see that an UP or DOWN has been performed on schedule, but will gain no more than one bit of metadata on every timer tick.

To implement this one-bit leakage, \project must either push up or pull down a message on each timer tick, whether or not it is has a message to deliver or a deletion scheduled. This condition is satisfied through artifically constructed messages. A constructed message with contents of a random size chosen from a distribution mirroring that of messages usually sent and received of the user is generated, encrypted and packaged as if it were a legitimate message, saved to a FAKE MESSAGES PSEUDO-mailbox, and sent to the server. On deletion, a message from FAKE MESSAGES is chosen at random and deleted in lieu of a legitimate message. These actions appear identical to the server, since the server will see only updates to the INDEX and the append or delete of a message with subject ENCRYPTED <UNIQUE SUBJECT> and encrypted contents.

Even if the user never deletes a message, the scheduler should still periodically return a DOWN bit, so that the contents of the server do not needlessly continuously grow at a high rate. In the case where deletes never occur, one could imagine a schedule of only pushes, timed such that is it similar to the actual receipt rate, say every 20 minutes. This is an unacceptable rate of checking for mail in a modern system, where periodic updates occur an order of magnitude faster. Additionally, since only one message is pushed every $t$ minutes, messages received in a batch of $k$ would be delayed at least $k*t$ minutes.

% Learning Update Schedules - Exponential backoff?

\label{mixing}
\subsection{Message Mixing}
The scheduling mechanism is necessary but not sufficient to hide message arrival metadata from the server. Assume a user never deletes messages, for simplicity. Assume also that the server is maliciously attempting to track metadata, and notes the upload time of each message. Consider two messages, R and F. R is a real message, and F a fake. Initially, the server cannot distinguish between the two messages. Since R is real, there is a probability $p=0$ that it will be deleted at any point in the future. The goal of the scheduling algorithm is to send messages to the server at an effective rate similar to the rate of actual message receipt, so the server does not grow needlessly. Then, the number of fake messages will remain about constant, $k$. So at each timer tick where a DOWN occurs, a fake message will have probability $p=1/k$ of being deleted. Therefore, the probability that a message is not deleted after $n$ DOWN ticks is $(1-1/k)^n$ (since it must not be deleted at all time ticks), which tends towards 0 as $n$ increases. Thus, a fake message's long-term probability of being deleted is 1. So, after a length of time has passed, any message that was uploaded to the server and has remained there for a long period of time is likely a real message, and the server can therefore merely observe the time it was uploaded.

@@

\subsubsection{Hiding Among Many}
We can imagine a solution to this problem where many more fake messages are added than real messages, driving down the probability that a fake message is chosen for deletion until it resembles the probability that a real message is chosen, 0. (Note that even in this case we will still desire scheduled deletions, to disguise deletions of real messages.) Assume for simplicity that fake messages are added at an amortized constant rate. We would like to calculate the probability that a fake message is not deleted after the $n$ DOWN ticks following its upload at time $t$. This is equal to the product of the probabilities that a fake message was not deleted at all ticks between times $t$ and $t+n$. We can compute this based on the probability of deleting a message at a particular time after $i$ deletion intervals. Assume that at each each interval, $c+1$ real messages and $f$ fake messages came in. Then the probability that a message was fake is $f/(c+f+1)$, and $c+f+1$ messages total came in each interval. But one message was deleted each interval, so $c+f$ net messages are added. So after $i$ intervals, $i(c+f)$ messages are in the server total, and an expected $i(c+f)f/(c+f+1)\approx if$ of those are fake. So the probability of a message being deleted at time $i$ given that it is fake is $1/(if)$, and the probability that it is not deleted is $1-1/(if)$. Then if the message was uploaded at time $t$, the probability that it was not deleted at time $n$ is $\displaystyle\prod_{i=t}^{i=t+n}(1-\frac{1}{if})$. This product tends towards 0 as $n$ tends towards infinity, but its slope drops off quickly as $f$ increases.

@@

In this case, we need an $f$ of 64 to achieve merely a $p=0.89$ that a fake message is not deleted (compare, of course, with $p=1$ for a real message under the assumption that no real messages are ever deleted).

Is this feasible? Certainly with no space constraints, it is, but we must also consider this solution under reasonable constraints. Gmail, a popular email provider, allots 17 GB of space to a user. Assume that a user wants to use this email address for 5 years (a conservative estimate), and that messages average 170 kB (this number is accurate for my personal email account). Then, 100,000 messages will fit into the server. Say that we are comfortable with having 32 fake messages uploaded per delete operation. Even discounting other messages received in the mail server, this only allows 3125 delete operations, which amounts to 1.7 delete operations per day over the five years. This amount to a delay of about 7 hours any time the user would like to delete a message \--- and this is not taking into account the space that real messages will use. While a user without space constraints could certainly choose this method, we will now describe a more feasible solution.

\subsubsection{Mixing}
We can achieve metadata security while constraining mail server growth to a rate roughly consistent with the number of received messages. To do this, \project implements a mixing mechanism whereby scheduled updates push and pull not only a single message, but multiple messages among which the true target message will be disguised.

On a scheduled timer tick with action UP, \project selects MIX NUMBER messages from all messages with an entry in the current INDEX, (those being the messages that exist in CRYPTOBLOBS). It then FETCHes those messages from the server, reassigns their UNIQUE SUBJECTs, reuploads the new messages to the server, then deletes the messages that were fetched. When the messages are reuploaded, the new message is randomly inserted in with them, indistinguishable from the other messages. On a DOWN operation, a similar mixing occurs. MIX NUMBER + 1 messages are selected at random. If the message to be deleted is not among them, one of the selected messages is replaced with the message to actually be deleted. \project then fetches all MIX NUMBER + 1 messages, reassigns UNIQUE SUBJECTs, reuploads the MIX NUMBER messages that are not actually marked for deletion, then deletes all chosen messages. If there are fewer than MIX NUMBER messages to choose from, all existing messages are mixed.

>>>>

If a message has been received at the client from a secure source as in Section~\ref{mail-from-secure}, we would like to upload it to the IMAP server so that other \project clients can view it, and to achieve backwards compatibility with respect to user-facing functionality. Yet if we push a message to the server immediately upon its receipt, we are leaking the time of message receipt to the server. Thus, we need additional mechanisms for ensuring that a server that stores messages cannot determine when a message arrived.

@@

  \draw [arrow] (\project.10) -- node [above] {State information} (scheduler.-190);
  \draw [arrow] (scheduler.190) -- node [below] {Next action type} (\project.-10);

@@

This scheduler receives information from the \project module about its state. Namely, it is informed when a message is received at the client or marked for deletion on the server, thus is informed about the total number of messages on the server. Instead of immediately performing this action, the \project module will wait until the next timer tick, when it is called to perform its next scheduled action by its client. At that time, it asks the scheduler whether its next action is of type UP, DOWN, or NONE, where UP means that the server will grow in size by one message, DOWN means that it will shrink by one message, and NONE means that the total server size will remain constant. The scheduler considers how often messages are being received and how often scheduled communications happen, and returns a result based on its mode of action and the desired overall server growth rate relative to actual message receipt and deletion rates. The server will be able to see that an UP, DOWN, or NONE has been performed on schedule, but will gain no more than one trit of metadata on every timer tick.

To implement this one-trit leakage, \project must perform the determined action on each timer tick, whether or not it is has a message to deliver or a deletion scheduled. This condition is satisfied through artifically constructed messages. A constructed message with contents of a random size chosen from a distribution mirroring that of messages usually sent and received of the user is generated, encrypted and packaged as if it were a legitimate message, saved to a FAKE MESSAGES PSEUDO-mailbox, and sent to the server. On deletion, a message from FAKE MESSAGES is chosen at random and deleted in lieu of a legitimate message. These actions appear identical to a server that is not doing any message tracking, since the server will see only updates to the INDEX and the append or delete of a message with subject ENCRYPTED <UNIQUE SUBJECT> and encrypted contents.

\label{mixing}
\subsection{Message Mixing}

\subsubsection{Motivation}
A scheduling mechanism is not sufficient if we wish to allow message deletion or have a reasonable rate of server growth. As to the former, if a user deletes a message, then the server will know that that message was real, and will then know when the message arrived, as well. As to the rate of server growth, since \project can only send messages on a timed exchange and must send a message at all timed exchanges, the fake messages that \project is sending up will quickly fill the mail server if the client wishes to synchronize with the server at any reasonable rate.

To illustrate precisely how infeasible the latter suggestion is, consider the follow graph relating message check frequency for varying numbers of clients connected to an IMAP server and time to fill a mailbox, assuming that the server receives no real message. The total mailbox size is Gmail's current size of 17 GB, and an average message is 170 kB.

\includegraphics[width=\textwidth]{server_longevity}

With only two clients, if we wish to check our mail more than once every half hour, we can only use that server for three years - and this is assuming we do not receive any actual messages in that time. Of course, if we assume the ability to delete emails is a feature we want (and since our goal is full user-facing backwards compatibility, that is a reasonable assumption to make), then the longevity is irrelevant.

Nor could we simply delete fake messages periodically. While doing so would not immediately leak metadata as deleting real messages would, we could not delete enough messages to stunt growth while masking real messages among fake ones.

Assume in the first case that we would like server growth to roughly mirror the rate of incoming real messages. Consider two messages, R and F. R is a real message, and F a fake. Initially, the server cannot distinguish between the two messages. Since R is real, there is a probability $p=0$ that it will be deleted at any point in the future. The goal of the scheduling algorithm is to send messages to the server at an effective rate similar to the rate of actual message receipt, so the server does not grow needlessly. Then, the number of fake messages will remain about constant, $k$. So at each timer tick where a DOWN occurs, a fake message will have probability $p=1/k$ of being deleted. Therefore, the probability that a message is not deleted after $n$ DOWN ticks is $(1-1/k)^n$ (since it must not be deleted at all time ticks), which tends towards 0 as $n$ increases. Thus, a fake message's long-term probability of being deleted is 1. So, after a length of time has passed, any message that was uploaded to the server and has remained there for a long period of time is likely a real message, and the server can therefore merely observe the time it was uploaded.

@@

Now assume instead that we will let the server grow to some extent. By adding more fake messages than real messages, we will drive down the probability that a fake message is chosen for deletion until it resembles the probability that a real message is chosen, 0. How often could we delete messages chosen from the set of all fake messages at random before we lose the ability to hide real messages among the fake ones? Assume for simplicity that fake messages are added at an amortized constant rate. We would like to calculate the probability that a fake message is not deleted after the $n$ DOWN ticks following its upload at time $t$. This is equal to the product of the probabilities that a fake message was not deleted at all ticks between times $t$ and $t+n$. We can compute this based on the probability of deleting a message at a particular time after $i$ deletion intervals. Assume that at each each interval, $c+1$ real messages and $f$ fake messages came in. Then the probability that a message was fake is $f/(c+f+1)$, and $c+f+1$ messages total came in each interval. But one message was deleted each interval, so $c+f$ net messages are added. So after $i$ intervals, $i(c+f)$ messages are in the server total, and an expected $i(c+f)f/(c+f+1)\approx if$ of those are fake. So the probability of a message being deleted at time $i$ given that it is fake is $1/(if)$, and the probability that it is not deleted is $1-1/(if)$. Then if the message was uploaded at time $t$, the probability that it was not deleted at time $n$ is $\displaystyle\prod_{i=t}^{i=t+n}(1-\frac{1}{if})$. This product tends towards 0 as $n$ tends towards infinity, but its slope drops off quickly as $f$ increases.

@@

In this case, we need an $f$ of 64 to achieve merely a $p=0.89$ that a fake message is not deleted (compare, of course, with $p=1$ for a real message under the assumption that no real messages are ever deleted). Thus, to achieve even a relatively feasible probability match, we could delete only 1 in 64 messages, which does not particularly aid in reducing overall size.

\label{natural}
\subsubsection{Mixing with natural growth rate}

We can achieve metadata security while constraining mail server growth to a rate roughly consistent with the number of received messages. To do this, \project implements a mixing mechanism whereby scheduled updates push and pull not only a single message, but multiple messages among which the true target message will be disguised.

On a scheduled timer tick with action UP, \project selects MIX NUMBER messages from all messages with an entry in the current INDEX, (those being the messages that exist in CRYPTOBLOBS). It then FETCHes those messages from the server, reassigns their UNIQUE SUBJECTs, reuploads the new messages to the server, then deletes the messages that were fetched. When the messages are reuploaded, the new message is randomly inserted in with them, indistinguishable from the other messages. On a DOWN operation, a similar mixing occurs. MIX NUMBER + 1 messages are selected at random. If the message to be deleted is not among them, one of the selected messages is replaced with the message to actually be deleted. \project then fetches all MIX NUMBER + 1 messages, reassigns UNIQUE SUBJECTs, reuploads the MIX NUMBER messages that are not actually marked for deletion, then deletes all chosen messages. On a NONE operation, MIX NUMBER messages are pulled down, reassigned, and pushed back up. If there are fewer than MIX NUMBER messages to choose from, all existing messages are mixed.

To achieve a growth rate consistent with the rate of incoming messages, we would need to perform an infeasible amount of mixing between UP and DOWN operations. Intuitively, this is because to grow with the actual rate, each fake message that was in an UP operation must appear in a DOWN operation. Assume that we do not seed the mailbox with extraneous fake messages. Consider again messages R and F. This time, R and F were each one of MIX NUMBER + 1 messages pushed to the server. Without any further mixing, if the DOWN operation chooses exactly one message from the $2($MIX NUMBER $+ 1)$ messages that were pushed UP, then we know that the real message was in the other set of MIX NUMBER + 1 messages, and so immediately lose the benefit of having pushed up a fake message. Assuming MIX NUMBER $<<$ the total number of messages in the server, this will almost certainly be the case given that we must delete something.

We could mitigate this leak by mostly or completing mixing the contents of the server between operations. How long does it take to mix the contents? Assume that of $N$ total messages, we seed the server with $k$ fake messages, with a mix number of $m$. After $k$ steps, we have mixed the fake messages among $km$ (TODO: Finish doing the math for how long it takes to completely mix it, not sure the existing math is correct.)

\subsubsection{Mixing with accelerated growth rate}
The information leak in Section~\ref{natural} stems from the difficulty of decorrelating UP and DOWN operations. A natural solution is to remove these operations entirely, using only NONE operations. We could implement this by seeding the server with an adequate number of fake messages, completely mixing it, and implementing message pushes as replacing a fake message with a real message, and message deletions as replacing a real message with a fake message. Then, UP operations could be solely for the purpose of keeping the number of fake messages in an expanding mailbox up to a reasonable percentage. With an adequate percentage of fake messages, we could choose to avoid the Monty Hall problem of identification by future selection for deletion by choosing entirely at random, which will still give reasonable rates of viable message insertion vectors.

(TODO: do math here.)
